---
title: "R Notebook"
---

# Combine files 

In this notebook, we are going to generate a helper function that takes in a few files and combines them into one. 

The two requirements to use this function are: 
1. the files have to have the same data columns 
2. the files have to have a common part of the file name, to identify them with 

# Helper function 

```{r}
combine_files <- function(location, filepattern, outputfilename){
  found_files <- list.files(path=location, pattern=filepattern, full.names = TRUE)
  output_df <- do.call(rbind, lapply(found_files, function(x) fread(file=x) ) )
  #fwrite(output_df, file = outputfilename)
  return(found_files)
}
```


# Application

I created this function in order to piece together SN 49 raw files, which I had to download separately. I will apply the function below to this example: 


First, we will define the variables

Remember that this notebook is in the 'files' subfolder, so we will have to go one folder up in order to access the data folder, or save to different locations 

```{r}
location = "./data/quant_aq_March2021"
filepattern = "sn49_raw_\\d.csv" #find all files with "sn49_raw_" followed by a number followed by csv
outputfilename = "./data/quant_aq_March2021/sn49_raw.csv"
```

We will have to call all the libraries we are using: 
```{r}
library(data.table)
```

Now we will run the function 

```{r}
output_df1<- combine_files(location, filepattern, outputfilename)
```

