---
title: "Combine QuantAQ files"
---

# Importing Libraries 

```{r, warning= FALSE, error = FALSE}
library(lubridate) #date and time functions 
library(data.table) #to use the data.table variable type
library(dplyr) #this library allows you to use the %>% operator
library(tidyr) #this library lets you use the complete function to account for time syncing
library(openair) #for plotting and analysis
library(stringr)
library(baseline) # for baseline correction
```

# Creating Functions 

```{r}
clean_sensor_data <- function(sensor){
  sensor$date <- ymd_hms(sensor$timestamp, tz="UTC") #parse datetime
  #sensor$date <- with_tz(sensor$date, "America/New_York") #change the time according to shifted timezone
  sensor <- mutate(sensor, originaldate = date) #keeping original times for comparing to flight data
  sensor$date <- round_date(ymd_hms(sensor$date, tz="UTC"), unit="minute") #round date for merging
  sensor<- sensor[order(sensor$originaldate),] #put it in chronological order
  return(sensor)
}
```

# Importing the Box Files 

```{r}
# #defining which variables we'll keep
box_vars <- c( "timestamp","timestamp_local","temp_box",  "temp_manifold", "rh_manifold", "pressure",
                  "noise", "solar", "wind_dir", "wind_speed", "co", "no" , "no2", "o3" ,  "pm1" , "pm25",
                  "pm10", "co2" ,"bin0", "bin1", "bin2", "bin3", "bin4", "bin5", "no_ae", "co_ae", "no2_ae") #raw variables to keep
```

```{r}
temp = list.files(path = "../data/boxdata/", pattern = "^.*-f0.*\\.csv", full.names = TRUE) #specify files we're looking for
boxfiles = sapply(temp,  FUN = function(x) fread(file=x,  data.table = TRUE, select = box_vars), simplify = FALSE,USE.NAMES = TRUE) 
```

```{r}
new_names <- lapply(names(boxfiles), function(x) substr(x, 17, 26)) #extract the real sensor names from the current names used in the list
names(boxfiles) <- new_names #rename list names
names(boxfiles)
```

```{r}
names(boxfiles) <- lapply(names(boxfiles), function(x) str_replace(x, "Output-0", "sn"))
names(boxfiles)
```

```{r}
boxfiles <- lapply(boxfiles, function(x) clean_sensor_data(x))
```


# Importing the QuantAQ Files 

```{r}
# #defining which variables we'll keep
final_vars <- c( "timestamp","timestamp_local","temp_box",  "temp_manifold", "rh_manifold", "pressure",
                "noise", "solar", "wind_dir", "wind_speed", "co", "no" , "no2", "o3" ,  "pm1" , "pm25",
                "pm10", "co2" ) #final variables to keep

# temp_box	temp_manifold	rh_manifold	pressure	noise	solar	wind_dir	wind_speed
raw_vars <- c("timestamp","bin0", "bin1", "bin2", "bin3", "bin4", "bin5", "no_ae", "co_ae", "no2_ae") #raw variables to keep
```

```{r}
#finding all final datasets
temp2 = list.files(path = "../data/quantAQwebsitedata/", pattern = "^.*_final.*\\.csv", full.names = TRUE)
finalfiles = sapply(temp2,  FUN = function(x) fread(file=x,  data.table = TRUE, select = final_vars), simplify = FALSE,USE.NAMES = TRUE)
#finalfiles = sapply(temp, function(x) fread(file=x , data.table = TRUE) , simplify = FALSE,USE.NAMES = TRUE)

#finding all raw datasets
temp3 = list.files(path = "../data/quantAQwebsitedata/", pattern = "^.*_raw.*\\.csv", full.names = TRUE)
rawfiles = sapply(temp3, function(x) fread(file=x, data.table = TRUE, select = raw_vars) , simplify = FALSE,USE.NAMES = TRUE)
#rawfiles = sapply(temp, function(x) fread(file=x,  data.table = TRUE) , simplify = FALSE,USE.NAMES = TRUE)
```

```{r}
names(finalfiles)<- lapply(names(finalfiles), function(x) substr(x, 10+18, 10+21)) #extract the real sensor names from the current names used in the list
names(rawfiles)<- lapply(names(rawfiles), function(x) substr(x, 10+18, 11+20))
names(finalfiles)
names(rawfiles)
``` 

```{r}
rawfiles <- lapply(rawfiles, function(x) clean_sensor_data(x))
rawfiles <- lapply(rawfiles, function(x) unique(x)) 
finalfiles <- lapply(finalfiles, function(x) clean_sensor_data(x))
finalfiles <- lapply(finalfiles, function(x) unique(x)) 
```

```{r}
quantAQfiles <- sapply(names(rawfiles), function(k) merge.data.table(finalfiles[[k]], rawfiles[[k]], by="date", all.x = TRUE), simplify = FALSE,USE.NAMES = TRUE)
```

# Finding the end date of the Box Files 

```{r}
enddates <- lapply(boxfiles, function(x) tail(x$date, n=1))
```

# Cutting the QuantAQ data at the end date 

```{r}
cut_quantAQfiles <- sapply(names(quantAQfiles), function(x) subset(quantAQfiles[[x]], date> enddates[[x]]), simplify = FALSE,USE.NAMES = TRUE )
```

# Joining the two datasets


```{r}
sn45joined <- rbind(boxfiles$sn45, cut_quantAQfiles$sn45, fill = TRUE)
sn46joined <- rbind(boxfiles$sn46, cut_quantAQfiles$sn46, fill = TRUE)
sn49joined <- rbind(boxfiles$sn49, cut_quantAQfiles$sn49, fill = TRUE)
sn62joined <- rbind(boxfiles$sn62, cut_quantAQfiles$sn62, fill = TRUE)
sn67joined <- rbind(boxfiles$sn67, cut_quantAQfiles$sn67, fill = TRUE)
sn72joined <- rbind(boxfiles$sn72, cut_quantAQfiles$sn72, fill = TRUE)

joined_files <- list("sn45" = sn45joined, "sn46" = sn46joined, "sn49" = sn49joined, "sn62"= sn62joined, "sn67" = sn67joined, "sn72" = sn72joined)

```


```{r}
mapply(
  fwrite, #apply function write table
  x=joined_files, file=paste("../data/joined/", names(joined_files), ".csv", sep = ""), #for each dataframe, use its name to make a csv of it
  MoreArgs=list(row.names=FALSE, sep=",")
)
```


```{r}
# lapply(joined_files, function(x) fread(x, file = paste(x, "csv", sep=".")))
```

# Combining SN 49 raw files 

Import all the files and rbind them together as they are being imported 

```{r}
# sn49files <- list.files(path = "./data/sn49_raw/", pattern = "*.csv", full.names = TRUE)
# sn49raw <-  do.call(rbind, lapply(sn49files, function(x) fread(file=x) ) )
```

And put it in the right order:
```{r}
# sn49raw <- sn49raw[(order(ymd_hms(sn49raw$timestamp))),]
```

Use the unique function to remove duplicate rows 

```{r}
# sn49raw <- unique(sn49raw)
```

Save the resulting dataframe to a file

```{r}
# fwrite(sn49raw, file = "./data/quantAQwebsitedata/sn49_raw.csv")
```

