---
title: "City Hall HEPA Data Analysis"
author: "Megan Ku"
date: "3/4/2021"
output: html_document
---

```{r setup, warning= FALSE, error = FALSE}
library(lubridate) #date and time functions 
library(data.table) #to use the data.table variable type
library(dplyr) #this library allows you to use the %>% operator
library(tidyr) #this library lets you use the complete function to account for time syncing
library(openair) #for plotting and analysis
library(stringr)
library(readr) #read .csv files
library(ggplot2) #plotting
source("filter_data.R")
source("make_boxplots.R")
```

# HEPA Impact Plots for Revere City Hall and Revere VFW

This notebook focuses on creating visuals to highlight the impact of HEPA
purifiers on particle concentration in classrooms.

## Setup

### Import Data

We will start by importing the datasets here. 

The variables of interest to us are the following:

- concent_ratio: indoor/outdoor concentration ratio for CPC data as calculated 
in hepa_main.Rmd
- concent_1.indoor/.outdoor: used to calculate concent_ratio
- bin0: smallest particle size from Mod-PM
- neph_bin0: used to calculate pm1, pm2.5, and pm10 (nephlometer)
- pm1, pm2.5, pm10: regulated particle sizes from EPA

We can ignore parsing failures for columns we won't be utilizing. We'll combine 
the imported data frames into one list.

```{r}
# Load data
auditor_ratio <- import_ratio_data('./data/CH_auditor_filt.csv', TRUE)
mainroom_ratio <- import_ratio_data('./data/CH_mainroom_filt.csv', TRUE)
waterdept_ratio <- import_ratio_data('./data/CH_waterdept_filt.csv', TRUE)

clerk_ratio <- import_ratio_data('./data/CH_clerk_filt.csv')
engineering_ratio <- import_ratio_data('./data/CH_engineering_filt.csv')
veterans_ratio <- import_ratio_data('./data/CH_veterans_filt.csv')
```

```{r}
# Create list
CPC_ratio_list <- list(
  "auditor" = auditor_ratio,
  "mainroom" = mainroom_ratio,
  "waterdept" = waterdept_ratio
)
PM_ratio_list <- list(
  "clerk" = clerk_ratio,
  "engineering" = engineering_ratio,
  "veterans" = veterans_ratio
)
```

### Isolate columns

To make the data frame more manageable, we'll only select specific columns. We 
will also calculate indoor/outdoor ratios for all particle sizes. concent_ratio
and pm1_ratio were already calculated in hepa_main.Rmd.

```{r}
CPC_focus_list <- lapply(CPC_ratio_list,  simplify_ratio_data, is_CPC=TRUE)
PM_focus_list <- lapply(PM_ratio_list,  simplify_ratio_data) 
```
## Pre-Processing

### Tidy dataset

It is good practice that a dataset be *tidy*. We will make this data tidier by
transposing the columns.

```{r}
CPC_focus_tidy <- lapply(CPC_focus_list,  function(x) tidy_data(x))
PM_focus_tidy <- lapply(PM_focus_list, function(x) tidy_data(x))
```

### Identify pre/post HEPA purifiers

Let's filter the data for the times before and after the HEPA purifiers were
installed.

```{r}
# Create new list to store dataframes
CPC_tidy_split <- vector(mode = "list", length = 1)
PM_tidy_split <- vector(mode = "list", length = 1)

CPC_tidy_split$auditor.post <- CPC_focus_tidy$auditor %>%
  filter(date1min > as.POSIXct("2021-01-19 09:20:00", tz='America/New_York'))
CPC_tidy_split$auditor.pre <- CPC_focus_tidy$auditor %>%
  filter(date1min < as.POSIXct("2021-01-19 09:20:00", tz='America/New_York'))

CPC_tidy_split$mainroom.post <- CPC_focus_tidy$mainroom %>%
  filter(date1min > as.POSIXct("2021-01-27 12:30:00", tz='America/New_York'))
CPC_tidy_split$mainroom.pre <- CPC_focus_tidy$mainroom %>%
  filter(date1min < as.POSIXct("2021-01-27 12:30:00", tz='America/New_York'))

CPC_tidy_split$waterdept.post <- CPC_focus_tidy$waterdept %>%
  filter(date1min > as.POSIXct("2021-01-11 10:05:00", tz='America/New_York'))
CPC_tidy_split$waterdept.pre <- CPC_focus_tidy$waterdept %>%
  filter(date1min < as.POSIXct("2021-01-11 10:05:00", tz='America/New_York'))


PM_tidy_split$engineering.post <- PM_focus_tidy$engineering %>%
  filter(date1min > as.POSIXct("2021-01-11 10:45:00", tz='America/New_York'))
PM_tidy_split$engineering.pre <- PM_focus_tidy$engineering %>%
  filter(date1min < as.POSIXct("2021-01-11 10:45:00", tz='America/New_York'))

PM_tidy_split$veterans.post <- PM_focus_tidy$veterans %>%
  filter(date1min > as.POSIXct("2021-01-27 09:30:00", tz='America/New_York'))
PM_tidy_split$veterans.pre <- PM_focus_tidy$veterans %>%
  filter(date1min < as.POSIXct("2021-01-27 09:30:00", tz='America/New_York'))

PM_tidy_split$clerk.post <- PM_focus_tidy$clerk %>%
  filter(date1min > as.POSIXct("2021-01-19 09:11:00", tz='America/New_York'))
PM_tidy_split$clerk.pre <- PM_focus_tidy$clerk %>%
  filter(date1min < as.POSIXct("2021-01-19 09:11:00", tz='America/New_York'))

# Remove placeholders
CPC_tidy_split <- CPC_tidy_split[lengths(CPC_tidy_split) != 0]
PM_tidy_split <- PM_tidy_split[lengths(PM_tidy_split) != 0]
```

### Work Hours (TODO: Determine if this is necessary or omit completely)

We will now filter the data for work hours to focus on the impact the purifiers
have on the people in these buildings. For City Hall and VFW, let's say that 
normal work hours are M-F from 8am to 5pm ET.

```{r}
# ratio_list_workday <- lapply(ratio_list,  function(x) filter_workday(x))
```

## Mean Concentrations

Let's calculate the mean concentration ratios before and after the HEPA 
installation, as well as the standard deviations. We'll store this in another 
dataset, which we will also tidy up.

```{r}
names = c("auditor", "waterdept", "mainroom", "auditor", "waterdept", "mainroom")

is_HEPA = c("pre", "pre", "pre", "post", "post", "post")


CPC_mean_ratios <- lapply(CPC_tidy_split, calculate_mean_ratio, TRUE) %>%
  bind_rows()

# Add labels to dataframe
CPC_mean_ratios <- cbind(name = names, CPC_mean_ratios)
CPC_mean_ratios <- cbind(CPC_mean_ratios, is_HEPA)

# Pivot wider for easier readability upon export
CPC_mean_ratios <- CPC_mean_ratios %>%
  pivot_wider(names_from = is_HEPA, names_sep = '.', values_from = -name) %>%
  mutate(conc_reduction = (conc.pre - conc.post)/conc.pre,
         bin0_reduction = (bin0.pre - bin0.post)/bin0.pre,
         neph_bin0_reduction = (neph_bin0.pre - neph_bin0.post)/neph_bin0.pre,
         pm1_reduction = (pm1.pre - pm1.post)/pm1.pre,
         pm25_reduction = (pm25.pre - pm25.post)/pm25.pre,
         pm10_reduction = (pm10.pre - pm10.post)/pm10.pre)

CPC_mean_ratios

#PM-only data
PM_mean_ratios <- lapply(PM_tidy_split, calculate_mean_ratio) %>%
  bind_rows() 

names = c("engineering", "veterans", "clerk", "engineering", "veterans", "clerk")
is_HEPA = c("pre", "pre", "pre", "post", "post", "post")

PM_mean_ratios <- cbind(name = names, PM_mean_ratios)
PM_mean_ratios <- cbind(PM_mean_ratios, is_HEPA)

PM_mean_ratios <- PM_mean_ratios %>%
  pivot_wider(names_from = is_HEPA, names_sep = '.', values_from = -name) %>%
  mutate(bin0_reduction = (bin0.pre - bin0.post)/bin0.pre,
         neph_bin0_reduction = (neph_bin0.pre - neph_bin0.post)/neph_bin0.pre,
         pm1_reduction = (pm1.pre - pm1.post)/pm1.pre,
         pm25_reduction = (pm25.pre - pm25.post)/pm25.pre,
         pm10_reduction = (pm10.pre - pm10.post)/pm10.pre)

PM_mean_ratios

```
Here, we'll export the data to a .csv file for further data analysis.

```{r}
write.table(CPC_mean_ratios, "./data/CH_mean_ratios_CPC.csv", append = FALSE, sep = ",",
            dec = ".", row.names = FALSE, col.names = TRUE)
write.table(PM_mean_ratios, "./data/CH_mean_ratios_PM.csv", append = FALSE, sep = ",",
            dec = ".", row.names = FALSE, col.names = TRUE)
```


## Plot Concentration Changes

### HEPA data tag

Let's investigate the data by using boxplots. Before we do that, though, we 
must distinguish which data is before and after the purifier installation.

```{r}
# Combine Water Department data
CPC_tidy_split$waterdept.pre <- CPC_tidy_split$waterdept.pre %>%
  mutate(is_HEPA = "no")

CPC_tidy_split$waterdept.post <- CPC_tidy_split$waterdept.post %>%
  mutate(is_HEPA = "yes")

waterdept_combined <- rbind(CPC_tidy_split$waterdept.pre, CPC_tidy_split$waterdept.post)


# Prep for boxplot
waterdept_boxplot_ratios <- waterdept_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

waterdept_boxplot_ratios$source <- factor(waterdept_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Auditor data
CPC_tidy_split$auditor.pre <- CPC_tidy_split$auditor.pre %>%
  mutate(is_HEPA = "no")

CPC_tidy_split$auditor.post <- CPC_tidy_split$auditor.post %>%
  mutate(is_HEPA = "yes")

auditor_combined <- rbind(CPC_tidy_split$auditor.pre, CPC_tidy_split$auditor.post)

auditor_boxplot_ratios <- auditor_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

# Prep for boxplot
auditor_boxplot_ratios$source <- factor(auditor_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Main Room data
CPC_tidy_split$mainroom.pre <- CPC_tidy_split$mainroom.pre %>%
  mutate(is_HEPA = "no")

CPC_tidy_split$mainroom.post <- CPC_tidy_split$mainroom.post %>%
  mutate(is_HEPA = "yes")

mainroom_combined <- rbind(CPC_tidy_split$mainroom.pre, CPC_tidy_split$mainroom.post)

mainroom_boxplot_ratios <- mainroom_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

# Prep for boxplot
mainroom_boxplot_ratios$source <- factor(mainroom_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))
```

We will use the built-in ggplot2 boxplot and adjust the visuals to our needs.

### Water Department Room

```{r, fig.height=5}

# Nice labels for facet_grid
source.labs <- c("CPC Conc.", "Bin 0", "Neph. Bin 0", "PM1", "PM2.5", "PM10")
names(source.labs) <- c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio")

waterdept_boxplot <- ggplot(data=waterdept_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +

  # separate visually by measurement and move labels to bottom of boxplot
  facet_grid(. ~ source, scales='free_x',
             labeller = labeller(source = source.labs),
             switch = "x") +
  
  # label graph
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Water Department Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  # Change legend and use blue theme
  scale_fill_brewer(name = "Sensor Measurements",
                    labels = c("Initial Recording", 
                                 "After HEPA Installation"),
                    palette="Blues") +
  

  # Clean up axes
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +

  # Set limits  
  ylim(0,1)

waterdept_boxplot

# Save image
ggsave("waterdept_box.png")
```

### Auditor Room

```{r, fig.height=5}
auditor_boxplot <- ggplot(data=auditor_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  # separate visually by measurement and move labels to bottom of boxplot
  facet_grid(. ~ source, scales='free_x', 
             labeller = labeller(source = source.labs),
             switch = "x") +
  

  # label graph
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Auditor Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  # Change legend and use blue theme
  scale_fill_brewer(name = "Sensor Measurements",
                  labels = c("Initial Recording", 
                               "After HEPA Installation"),
                  palette="Blues") +
  
  # Clean up axes
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  
  # Set limits
  ylim(0,1)

auditor_boxplot

# Save image
ggsave("auditor_box.png")
```
### VFW Main Room

```{r, fig.height=5}
mainroom_boxplot <- ggplot(data=mainroom_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  # separate visually by measurement and move labels to bottom of boxplot
  facet_grid(. ~ source, scales='free_x', 
             labeller = labeller(source = source.labs),
             switch = "x") +
  

  # label graph
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "VFW: Main Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  # Change legend and use blue theme
  scale_fill_brewer(name = "Sensor Measurements",
                  labels = c("Initial Recording", 
                               "After HEPA Installation"),
                  palette="Blues") +
  
  # Clean up axes
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  
  # Set limits
  ylim(0,1)

mainroom_boxplot

# Save image
ggsave("mainroom_box.png")
```

# PM-only Data Boxplots

```{r}
# Combine Veterans data
PM_tidy_split$veterans.pre <- PM_tidy_split$veterans.pre %>%
  mutate(is_HEPA = "no")

PM_tidy_split$veterans.post <- PM_tidy_split$veterans.post %>%
  mutate(is_HEPA = "yes")

veterans_combined <- rbind(PM_tidy_split$veterans.pre, PM_tidy_split$veterans.post)

veterans_boxplot_ratios <- veterans_combined %>%
  filter(source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

veterans_boxplot_ratios$source <- factor(veterans_boxplot_ratios$source , levels=c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Engineering data
PM_tidy_split$engineering.pre <- PM_tidy_split$engineering.pre %>%
  mutate(is_HEPA = "no")

PM_tidy_split$engineering.post <- PM_tidy_split$engineering.post %>%
  mutate(is_HEPA = "yes")

engineering_combined <- rbind(PM_tidy_split$engineering.pre, PM_tidy_split$engineering.post)

engineering_boxplot_ratios <- engineering_combined %>%
  filter(source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

engineering_boxplot_ratios$source <- factor(engineering_boxplot_ratios$source , levels=c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Clerk's Office data
PM_tidy_split$clerk.pre <- PM_tidy_split$clerk.pre %>%
  mutate(is_HEPA = "no")

PM_tidy_split$clerk.post <- PM_tidy_split$clerk.post %>%
  mutate(is_HEPA = "yes")

clerk_combined <- rbind(PM_tidy_split$clerk.pre, PM_tidy_split$clerk.post)

clerk_boxplot_ratios <- clerk_combined %>%
  filter(source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

clerk_boxplot_ratios$source <- factor(clerk_boxplot_ratios$source , levels=c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))
```


```{r, fig.height=5}
source.labs <- c("Bin 0", "Neph. Bin 0", "PM1", "PM2.5", "PM10")
names(source.labs) <- c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio")

veterans_boxplot <- ggplot(data=veterans_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x',
             labeller = labeller(source = source.labs),
             switch = "x") +
  
  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere VFW: Veteran Services Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  
  scale_fill_brewer(name = "Sensor Measurements",
                    labels = c("Initial Recording", 
                                 "After HEPA Installation"),
                    palette="Blues") +
  

  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +

  ylim(0,1)

veterans_boxplot
ggsave("veterans_box.png")
```

```{r, fig.height=5}
source.labs <- c("Bin 0", "Neph. Bin 0", "PM1", "PM2.5", "PM10")
names(source.labs) <- c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio")

engineering_boxplot <- ggplot(data=engineering_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x',
             labeller = labeller(source = source.labs),
             switch = "x") +
  
  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Engineering Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  
  scale_fill_brewer(name = "Sensor Measurements",
                    labels = c("Initial Recording", 
                                 "After HEPA Installation"),
                    palette="Blues") +
  

  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +

  ylim(0,1)

engineering_boxplot
ggsave("engineering_box.png")
```

```{r, fig.height=5}
source.labs <- c("Bin 0", "Neph. Bin 0", "PM1", "PM2.5", "PM10")
names(source.labs) <- c("bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio")

clerk_boxplot <- ggplot(data=clerk_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x',
             labeller = labeller(source = source.labs),
             switch = "x") +
  
  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Clerk's Office",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  
  scale_fill_brewer(name = "Sensor Measurements",
                    labels = c("Initial Recording", 
                                 "After HEPA Installation"),
                    palette="Blues") +
  

  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) + 

  ylim(0,1)

clerk_boxplot
ggsave("clerk_box.png")
```