---
title: "hepa_impact"
author: "Megan Ku"
date: "3/4/2021"
output: html_document
---

```{r setup, warning= FALSE, error = FALSE}
library(lubridate) #date and time functions 
library(data.table) #to use the data.table variable type
library(dplyr) #this library allows you to use the %>% operator
library(tidyr) #this library lets you use the complete function to account for time syncing
library(openair) #for plotting and analysis
library(stringr)
library(readr) #read .csv files
library(ggplot2) #plotting
```

## HEPA Impact Plot

This notebook focuses on creating visuals to highlight the impact of HEPA
purifiers on particle concentration in classrooms.

### Setup
In hepa_main.Rmd, un-comment the code in the chunk at line 486 to download the 
processed data as .csv files. 

We will start by importing them here. We need to manually set concent_ratio as 
a double.

```{r}
waterdept_ratio <- read_csv('waterdept_ratio.csv')

auditor_ratio <- read_csv('auditor_ratio.csv', col_types = cols(
  concent_ratio = col_double()
))

mainroom_ratio <- read_csv('mainroom_ratio.csv', col_types = cols(
  concent_ratio = col_double()
))

head(waterdept_ratio)
```

Looking at hepa_main.Rmd, it seems that it's more useful to keep all structures
as a list.

```{r}
ratio_list <- list(
  "auditor" = auditor_ratio,
  "waterdept" = waterdept_ratio,
  "mainroom" = mainroom_ratio
  # "engineering" = engineering_ratio
  )
```


To make the data frame more manageable, we'll only select these columns:
date1min, 
concent_ratio, 
bin0.outdoor & bin0.indoor, 
neph_bin0.outdoor, & neph_bin0.indoor, 
pm1_ratio

```{r}
simplify_ratio_data <- function(room) {
  room <- room %>%
    # Add/delete relevant columns here
    mutate(bin0_ratio = bin0.indoor/bin0.outdoor) %>%
    mutate(neph_bin0_ratio = neph_bin0.indoor/neph_bin0.outdoor) %>%
    mutate(pm25_ratio = pm25.indoor/pm25.outdoor) %>%
    mutate(pm10_ratio = pm10.indoor/pm10.outdoor) %>%
    select(date1min, 
           concent_ratio,
           concent_1.indoor,
           concent_1.outdoor,
           bin0_ratio,
           bin0.indoor,
           bin0.outdoor,
           neph_bin0_ratio,
           neph_bin0.indoor,
           neph_bin0.outdoor,
           pm1_ratio,
           pm1.indoor,
           pm1.outdoor,
           pm25_ratio,
           pm25.indoor,
           pm25.outdoor,
           pm10_ratio,
           pm10.indoor,
           pm10.outdoor)
    
  return(room)
}

ratio_list <- lapply(ratio_list,  function(x) simplify_ratio_data(x)) # Only save columns of importance 
```

Since times were imported as UTC, the below code block switches them back to
Eastern.

```{r}
time_to_eastern <- function(room) {
  tz(room$date1min) <- "America/New_York"
  return(room)
}

ratio_list <- lapply(ratio_list,  function(x) time_to_eastern(x)) # Fix time zone
```

```{r}
head(ratio_list$auditor)
```
It is good practice that a dataset be *tidy*. We will make this data tidier by
transposing the columns.

```{r}
tidy_data <- function(room) {
  room <- room %>%
    pivot_longer(-date1min, names_to = "source", values_to = "ratio")
}

ratio_list <- lapply(ratio_list,  function(x) tidy_data(x))

head(ratio_list$auditor)
```




We will now filter the data for work hours to focus on the impact the purifiers
have on the people in these buildings. For City Hall and VFW, let's say that 
normal work hours are M-F from 8am to 5pm ET.

```{r}
filter_workday <- function(room) {
  room <- room %>%
    filter(hour(date1min) >= 8 & hour(date1min) <= 17) %>%
    filter(wday(date1min) %in% 2:6)
}

ratio_list <- lapply(ratio_list,  function(x) filter_workday(x))

ratio_list$waterdept
```
The main room has a lot of data points that go to (nearly) zero, so we're going 
to delete those rows. I inspected the data and noticed that these points were
scattered among 0 readings.

```{r}
ratio_list$mainroom <- ratio_list$mainroom %>%
  subset(ratio > 0.0001)
```


Let's filter the data for the times before and after the HEPA purifiers were
installed. Dataframes with '_HEPA' indicate data taken post HEPA purifier 
installation.

```{r}

ratio_list$auditor_HEPA <- ratio_list$auditor %>%
  filter(date1min > as.POSIXct("2021-01-19 09:20:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-25 09:03:00", tz='America/New_York'))

ratio_list$mainroom_HEPA <- ratio_list$mainroom %>%
  filter(date1min > as.POSIXct("2021-01-27 12:30:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-27 19:53:00", tz='America/New_York'))

ratio_list$waterdept_HEPA <- ratio_list$waterdept %>%
  filter(date1min > as.POSIXct("2021-01-11 10:05:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-13 12:08:00", tz='America/New_York'))

ratio_list$auditor <- ratio_list$auditor %>%
  filter(date1min > as.POSIXct("2021-01-13 12:49:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-19 09:20:00", tz='America/New_York'))

ratio_list$mainroom <- ratio_list$mainroom %>%
  filter(date1min > as.POSIXct("2021-01-25 09:58:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-27 12:30:00", tz='America/New_York'))

ratio_list$waterdept <- ratio_list$waterdept %>%
  filter(date1min > as.POSIXct("2021-01-06 08:45:00", tz='America/New_York'), 
         date1min < as.POSIXct("2021-01-11 10:05:00", tz='America/New_York'))
```

Let's calculate the mean concentration ratios before and after the HEPA 
installation, as well as the standard deviations. We'll store this in another 
dataset, which we will also tidy up.

```{r}
names = c("auditor", "waterdept", "mainroom", "auditor", "waterdept", "mainroom")

is_HEPA = c("pre", "pre", "pre", "post", "post", "post")

calculate_mean_ratio <- function(room) {
  conc <- room %>%
    filter(source=="concent_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(conc = "mean(ratio, na.rm = TRUE)")
  conc_sd <- room %>%
    filter(source=="concent_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(conc_sd = "sd(ratio, na.rm = TRUE)")
  conc.indoor <- room %>%
    filter(source=="concent_1.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(conc.indoor = "mean(ratio, na.rm = TRUE)")
  conc_sd.indoor <- room %>%
    filter(source=="concent_1.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(conc_sd.indoor = "sd(ratio, na.rm = TRUE)")  
  conc.outdoor <- room %>%
    filter(source=="concent_1.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(conc.outdoor = "mean(ratio, na.rm = TRUE)")
  conc_sd.outdoor <- room %>%
    filter(source=="concent_1.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(conc_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  
  
  bin0 <- room %>%
    filter(source=="bin0_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(bin0 = "mean(ratio, na.rm = TRUE)")
  bin0_sd <- room %>%
    filter(source=="bin0_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(bin0_sd = "sd(ratio, na.rm = TRUE)")
  bin0.indoor <- room %>%
    filter(source=="bin0.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(bin0.indoor = "mean(ratio, na.rm = TRUE)")
  bin0_sd.indoor <- room %>%
    filter(source=="bin0.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(bin0_sd.indoor = "sd(ratio, na.rm = TRUE)")
    bin0.outdoor <- room %>%
    filter(source=="bin0.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(bin0.outdoor = "mean(ratio, na.rm = TRUE)")
  bin0_sd.outdoor <- room %>%
    filter(source=="bin0.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(bin0_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  
  neph_bin0 <- room %>%
    filter(source=="neph_bin0_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0 = "mean(ratio, na.rm = TRUE)")
  neph_bin0_sd <- room %>%
    filter(source=="neph_bin0_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0_sd = "sd(ratio, na.rm = TRUE)")
  neph_bin0.indoor <- room %>%
    filter(source=="neph_bin0.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0.indoor = "mean(ratio, na.rm = TRUE)")
  neph_bin0_sd.indoor <- room %>%
    filter(source=="neph_bin0.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0_sd.indoor = "sd(ratio, na.rm = TRUE)")
  neph_bin0.outdoor <- room %>%
    filter(source=="neph_bin0.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0.outdoor = "mean(ratio, na.rm = TRUE)")
  neph_bin0_sd.outdoor <- room %>%
    filter(source=="neph_bin0.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(neph_bin0_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  
  
  pm1 <- room %>%
    filter(source=="pm1_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm1 = "mean(ratio, na.rm = TRUE)")
  pm1_sd <- room %>%
    filter(source=="pm1_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm1_sd = "sd(ratio, na.rm = TRUE)")
  pm1.indoor <- room %>%
    filter(source=="pm1.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm1.indoor = "mean(ratio, na.rm = TRUE)")
  pm1_sd.indoor <- room %>%
    filter(source=="pm1.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm1_sd.indoor = "sd(ratio, na.rm = TRUE)")
  pm1.outdoor <- room %>%
    filter(source=="pm1.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm1.outdoor = "mean(ratio, na.rm = TRUE)")
  pm1_sd.outdoor <- room %>%
    filter(source=="pm1.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm1_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  
  
  pm25 <- room %>%
    filter(source=="pm25_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm25 = "mean(ratio, na.rm = TRUE)")
  pm25_sd <- room %>%
    filter(source=="pm25_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm25_sd = "sd(ratio, na.rm = TRUE)")
  pm25.indoor <- room %>%
    filter(source=="pm25.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm25.indoor = "mean(ratio, na.rm = TRUE)")
  pm25_sd.indoor <- room %>%
    filter(source=="pm25.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm25_sd.indoor = "sd(ratio, na.rm = TRUE)")
  pm25.outdoor <- room %>%
    filter(source=="pm25.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm25.outdoor = "mean(ratio, na.rm = TRUE)")
  pm25_sd.outdoor <- room %>%
    filter(source=="pm25.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm25_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  

  pm10 <- room %>%
    filter(source=="pm10_ratio") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm10 = "mean(ratio, na.rm = TRUE)")
  pm10_sd <- room %>%
    filter(source=="pm10_ratio") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm10_sd = "sd(ratio, na.rm = TRUE)")
  pm10.indoor <- room %>%
    filter(source=="pm10.indoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm10.indoor = "mean(ratio, na.rm = TRUE)")
  pm10_sd.indoor <- room %>%
    filter(source=="pm10.indoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm10_sd.indoor = "sd(ratio, na.rm = TRUE)")
  pm10.outdoor <- room %>%
    filter(source=="pm10.outdoor") %>%
    summarize(mean(ratio, na.rm=TRUE)) %>%
    rename(pm10.outdoor = "mean(ratio, na.rm = TRUE)")
  pm10_sd.outdoor <- room %>%
    filter(source=="pm10.outdoor") %>%
    summarize(sd(ratio, na.rm=TRUE)) %>%
    rename(pm10_sd.outdoor = "sd(ratio, na.rm = TRUE)")
  
  
  mean_ratios = cbind(conc, conc_sd, conc.indoor, conc_sd.indoor, conc.outdoor, conc_sd.outdoor,
                      bin0, bin0_sd, bin0.indoor, bin0_sd.indoor, bin0.outdoor, bin0_sd.outdoor,
                      neph_bin0, neph_bin0_sd, neph_bin0.indoor, neph_bin0_sd.indoor, neph_bin0.outdoor, neph_bin0_sd.outdoor, 
                      pm1, pm1_sd, pm1.indoor, pm1_sd.indoor, pm1.outdoor, pm1_sd.outdoor,
                      pm25, pm25_sd, pm25.indoor, pm25_sd.indoor, pm25.outdoor, pm25_sd.outdoor,
                      pm10, pm10_sd, pm10.indoor, pm10_sd.indoor, pm10.outdoor, pm10_sd.outdoor)
  return(mean_ratios)
}

mean_ratios <- lapply(ratio_list, function(x) calculate_mean_ratio(x)) %>%
  bind_rows() 

mean_ratios <- cbind(name = names, mean_ratios)
mean_ratios <- cbind(mean_ratios, is_HEPA)

mean_ratios <- mean_ratios %>%
  pivot_wider(names_from = is_HEPA, names_sep = '.', values_from = -name) %>%
  mutate(conc_reduction = (conc.pre - conc.post)/conc.pre,
         bin0_reduction = (bin0.pre - bin0.post)/bin0.pre,
         neph_bin0_reduction = (neph_bin0.pre - neph_bin0.post)/neph_bin0.pre,
         pm1_reduction = (pm1.pre - pm1.post)/pm1.pre,
         pm25_reduction = (pm25.pre - pm25.post)/pm25.pre,
         pm10_reduction = (pm10.pre - pm10.post)/pm10.pre)

mean_ratios

```
```{r}
write.table(mean_ratios, "revere_mean_ratios.csv", append = FALSE, sep = ",", dec = ".",
            row.names = TRUE, col.names = TRUE)
```


Let's investigate the data by using boxplots. Before we do that, though, we 
must distinguish which data is before and after the purifier installation.

```{r}
# Combine Water Department data
ratio_list$waterdept <- ratio_list$waterdept %>%
  mutate(is_HEPA = "no")

ratio_list$waterdept_HEPA <- ratio_list$waterdept_HEPA %>%
  mutate(is_HEPA = "yes")

waterdept_combined <- rbind(ratio_list$waterdept, ratio_list$waterdept_HEPA)

waterdept_boxplot_ratios <- waterdept_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

waterdept_boxplot_ratios$source <- factor(waterdept_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Auditor data
ratio_list$auditor <- ratio_list$auditor %>%
  mutate(is_HEPA = "no")

ratio_list$auditor_HEPA <- ratio_list$auditor_HEPA %>%
  mutate(is_HEPA = "yes")

auditor_combined <- rbind(ratio_list$auditor, ratio_list$auditor_HEPA)

auditor_boxplot_ratios <- auditor_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

auditor_boxplot_ratios$source <- factor(auditor_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))

# Combine Main Room data
ratio_list$mainroom <- ratio_list$mainroom %>%
  mutate(is_HEPA = "no")

ratio_list$mainroom_HEPA <- ratio_list$mainroom_HEPA %>%
  mutate(is_HEPA = "yes")

mainroom_combined <- rbind(ratio_list$mainroom, ratio_list$mainroom_HEPA)

mainroom_boxplot_ratios <- mainroom_combined %>%
  filter(source == "concent_ratio" | source == "bin0_ratio"| source == "neph_bin0_ratio" | source == "pm1_ratio" | source == "pm25_ratio" | source ==  "pm10_ratio")

mainroom_boxplot_ratios$source <- factor(mainroom_boxplot_ratios$source , levels=c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio"))
```

```{r}
waterdept_boxplot_ratios
```


```{r}
source.labs <- c("CPC Conc.", "Bin 0", "Neph. Bin 0", "PM1", "PM2.5", "PM10")
names(source.labs) <- c("concent_ratio", "bin0_ratio", "neph_bin0_ratio", 
                        "pm1_ratio", "pm25_ratio", "pm10_ratio")

waterdept_boxplot <- ggplot(data=waterdept_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x',
             labeller = labeller(source = source.labs),
             switch = "x") +
  
  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Water Department Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  
  scale_fill_brewer(name = "Sensor Measurements",
                    labels = c("Initial Recording", 
                                 "After HEPA Installation"),
                    palette="Blues") +
  

  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +

  ylim(0,1)

waterdept_boxplot
ggsave("waterdept_box.png")
```

```{r}
auditor_boxplot <- ggplot(data=auditor_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x', 
             labeller = labeller(source = source.labs),
             switch = "x") +
  

  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "Revere City Hall: Auditor Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  scale_fill_brewer(name = "Sensor Measurements",
                  labels = c("Initial Recording", 
                               "After HEPA Installation"),
                  palette="Blues") +
  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  
  ylim(0,1)

auditor_boxplot
ggsave("auditor_box.png")
```
```{r}
mainroom_boxplot <- ggplot(data=mainroom_boxplot_ratios, aes(x=source, y=ratio, fill=is_HEPA)) +
  geom_boxplot() +
  
  facet_grid(. ~ source, scales='free_x', 
             labeller = labeller(source = source.labs),
             switch = "x") +
  

  
  labs(title="Effect of HEPA Purifiers on Particle Ratios",
       subtitle= "VFW: Main Room",
       x="Particle Ratio Type",
       y="Indoor/Outdoor Ratio") +

  scale_fill_brewer(name = "Sensor Measurements",
                  labels = c("Initial Recording", 
                               "After HEPA Installation"),
                  palette="Blues") +
  
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  
  ylim(0,1)

mainroom_boxplot
ggsave("mainroom_box.png")
```

