---
title: "Data Check for ML Data"
---

# Import Data 

## Import Libraries 

```{r, warning= FALSE, error = FALSE}
library(lubridate) #date and time functions 
library(data.table) #to use the data.table variable type
library(dplyr) #this library allows you to use the %>% operator
library(tidyr) #this library lets you use the complete function to account for time syncing
library(openair) #for plotting and analysis
library(stringr)
library(baseline) # for baseline correction
```

## Import finalized data 

```{r}
nm2 <- list.files(path = "./data/final_whole/", pattern = "*.csv", full.names = TRUE)
snfiles <-  sapply(nm2,  FUN = function(x) fread(file=x,  data.table = TRUE), simplify = FALSE,USE.NAMES = TRUE)
```

```{r}
names(snfiles) <- c("sn45", "sn46", "sn49", "sn62", "sn67", "sn72")
```

```{r}
format_dates <- function(sensor){
  sensor$date <- ymd_hms(sensor$date, tz = "UTC")
  sensor$date_local <- ymd_hms(sensor$date_local, tz = "America/New_York")
  
  return(sensor)
}
```

```{r}
snfiles <- lapply(snfiles, function(x) format_dates(x))
```

## Import ML data 


```{r}
tempcsv <- list.files(path = "./data/MLdata/", pattern = "*.csv", full.names = TRUE)
MLfiles = sapply(tempcsv, function(x) fread(file=x, data.table = TRUE) , simplify = FALSE,USE.NAMES = TRUE) #import them in
names(MLfiles) <- c("sn45", "sn46", "sn49", "sn62", "sn67", "sn72")
```

```{r}
extractColumns <- function(sensor){
  sensor <- subset(sensor, select= c("timestamp", "co", "no", "no2", "o3", "flag" ))
}
```

```{r}
MLfiles <- lapply(MLfiles, function(x) extractColumns(x))
```

```{r}
clean_sensor_data <- function(sensor){
  sensor$date <- ymd_hms(sensor$timestamp, tz="UTC") #parse datetime
  #sensor$date <- with_tz(sensor$date, "America/New_York") #change the time according to shifted timezone
  sensor <- mutate(sensor, originaldate = date) #keeping original times for comparing to flight data
  sensor$date <- round_date(ymd_hms(sensor$date, tz="UTC"), unit="minute") #round date for merging
  sensor<- sensor[order(sensor$originaldate),] #put it in chronological order
  return(sensor)
}

get_localtime <- function(sensor){
  ##sensor$date_local = sensor$date
  sensor$date_local <- with_tz(sensor$date, tzone = "America/New_York")
  return(sensor)
}
```
```{r}
MLfiles <- lapply(MLfiles, function(x) clean_sensor_data(x))
MLfiles <- lapply(MLfiles, function(x) get_localtime(x))
```


## Merge data 

In order to compare the data, we have to merge the datasets 

```{r}
snfiles <- sapply(names(snfiles), function(k) merge.data.table(snfiles[[k]], MLfiles[[k]], by.x="date_local", by.y = "date_local", all = FALSE, suffixes = c("", ".ML")), simplify = FALSE,USE.NAMES = TRUE)
```

## Additional data cleaning 

we will add the data cleaning for the flags QuantAQ created to both the final and ML data. 

```{r}
flagFilter <- function(sensor){
  for(i in nrow(sensor)){
    if(sensor$flag[i] & 1 == 1 | sensor$flag[i] & 2 == 2){
      sensor$co[i] <- NA
      sensor$co.ML[i] <- NA
      sensor$no[i] <- NA
      sensor$no.ML[i] <- NA
      sensor$no2[i] <- NA
      sensor$no2.ML[i] <- NA
        }
    
    if(sensor$flag[i] & 4 == 4){
      sensor$co[i] <- NA
      sensor$co.ML[i] <- NA
    }
    
    if(sensor$flag[i] & 8 == 8){
      sensor$no[i] <- NA
      sensor$no.ML[i] <- NA
    }
    
    if(sensor$flag[i] & 16 == 16){
      sensor$no2[i] <- NA
      sensor$no2.ML[i] <- NA
    }
  }
  
  return(sensor)
}
```


```{r}
snfiles <- lapply(snfiles, function(x) flagFilter(x))
```

# Compare data 

## time series 

```{r}
pdf("ML-with-nofilters.pdf")
for (i in seq(length(names(snfiles)))){

  timePlot(snfiles[[i]], pollutant = c("no2", "no2.ML", "co",  "co.ML", "correctedNO",  "no.ML",  "o3", "o3.ML"),y.relation = "free", main = paste0(names(snfiles)[i], "Time Plots"),   cols = c("red", "green", "blue", "purple", "orange", "brown", "pink", "black"))
}

dev.off()
```

## correlation plot 

```{r}
for (i in seq(length(names(snfiles)))){

  corPlot(snfiles[[i]], pollutants = c("no2.ML", "no2", "co.ML", "co", "no.ML", "correctedNO",  "o3.ML", "o3"),y.relation = "free", main = paste0(names(snfiles)[i], "Correlation Plots"))
}
```


# Adding Quick Filters 

## Adding the filters 

We will apply the following quick filters to the ML data :

•	NO values higher than 1000 ppb on SN45
•	NO2 values higher than 300 ppb on SN45
•	All SN46 data after September 23, 2020 (recall that the sensor was moved)
•	Startup data on SN46 at the very beginning of the time series (Sept. 11, 2019 17:22-17:41)
•	NO values higher than 200 ppb on SN67
•	NO2 values higher than 80 ppb on SN67


```{r}
snfiles$sn45$no.ML[which(snfiles$sn45$no.ML > 1000)]  <- NA
snfiles$sn45$no2.ML[which(snfiles$sn45$no2.ML > 300)]  <- NA

snfiles$sn67$no.ML[which(snfiles$sn67$no.ML > 200)]  <- NA
snfiles$sn67$no2.ML[which(snfiles$sn67$no2.ML > 80)]  <- NA
  
snfiles$sn46 <- snfiles$sn46[which(date_local < "2020-09-23 00:00:00" & date_local > "2019-09-11 17:41:00")]
```

## Comparison 

```{r}
pdf("ML-with-quickfilters.pdf")
for (i in seq(length(names(snfiles)))){

  timePlot(snfiles[[i]], pollutant = c("no2", "no2.ML", "co",  "co.ML", "correctedNO",  "no.ML",  "o3", "o3.ML"),y.relation = "free", main = paste0(names(snfiles)[i], "Time Plots"),   cols = c("red", "green", "blue", "purple", "orange", "brown", "pink", "black"))
}

dev.off()
```

```{r}
# for (i in seq(length(names(snfiles)))){
# 
#   corPlot(snfiles[[i]], pollutants = c("no2.ML", "no2", "co.ML", "co", "no.ML", "correctedNO",  "o3.ML", "o3"),y.relation = "free", main = paste0(names(snfiles)[i], "Correlation Plots"))
# }
```


# Adding all data check filters 

## Adding the filters 

```{r}
removeCOspikes <- function(sensor){
  #calculate how much time elapses between data points
  sensor$timediff <- c(NA, difftime(sensor$date[-1],
                               sensor$date[-nrow(sensor)],
                               units="mins"))
  #find where there are jumps of an hour or more
  jump_indices <- which(sensor$timediff > 60)
  # set up which indices should be removed from the data 
  sensor$removeCO <- "FALSE"
  for(jump_index in jump_indices){
    sensor[jump_index:(jump_index+20), removeCO := "TRUE"] 
  }
  #remove CO 
  sensor$co.ML[sensor$removeCO == "TRUE"] <- NA 
  
  return(sensor)
}
```

```{r}
snfiles <- lapply(snfiles, function(x) removeCOspikes(x))
```


```{r}
apply_flags <- function(sensor){
  # creating a flags column to flag high values 
  # flag number to pollutant definition given below:
  # co - 1 
  # co2 - 2
  # no2 - 3
  # bin0 - 4
  # pm1 - 5
  # no - 6
  
  sensor$flag <- replicate(nrow(sensor), c(), simplify = FALSE) #create flags column which has a list for every data point
  sensor$flag[which(sensor$co.ML > 10000)] <- lapply(sensor$flag[which(sensor$co.ML > 10000)], function(x) c(x, 1)) #add 2 to the flags data points with high CO2 values

  sensor$flag[which(sensor$no2.ML > 1000)] <- lapply(sensor$flag[which(sensor$no2.ML > 1000)], function(x) c(x, 3))
  sensor$flag[which(sensor$no.ML > 300)] <- lapply(sensor$flag[which(sensor$no.ML > 300)], function(x) c(x, 6))

  return(sensor)
}
```

We can apply the flags to the sn dataframes: 

```{r}
snfiles <- lapply(snfiles, function(x) apply_flags(x)) 
```


```{r}
remove_flags <- function(sensor, flag_val, pollutant){
  sensor[which(sensor$flag %like% flag_val), pollutant] <- NA
  return(sensor)
}
```

```{r}
flags_list = list(c(1,"co.ML"), c(2,"co2"), c(3,"no2.ML"), c(6, "no.ML"))
for(i in flags_list){
  snfiles <- lapply(snfiles, function(x) remove_flags(x, i[1], i[2]))
}
```

```{r}
snfiles$sn45$flag <- NULL 
snfiles$sn46$flag <- NULL 
snfiles$sn49$flag <- NULL 
snfiles$sn62$flag <- NULL 
snfiles$sn67$flag <- NULL 
snfiles$sn72$flag <- NULL 
```


```{r}
negative_value_filter <- function(sensor){
  sensor$o3.ML <- replace(sensor$o3.ML, sensor$o3.ML < 0, 0)
  sensor$co.ML <- replace(sensor$co.ML, sensor$co.ML <0, NA)
  sensor$no2.ML <- replace(sensor$no2.ML, sensor$no2.ML <0, NA)
  return(sensor)
}
```

We'll apply this function to all the datasets:

```{r}
snfiles <- lapply(snfiles, function(x) negative_value_filter(x))
```


## Comparison 

```{r}
pdf("ML-with-allfilters.pdf")
for (i in seq(length(names(snfiles)))){

  timePlot(snfiles[[i]], pollutant = c("no2", "no2.ML", "co",  "co.ML", "correctedNO",  "no.ML",  "o3", "o3.ML"),y.relation = "free", main = paste0(names(snfiles)[i], "Time Plots"),   cols = c("red", "green", "blue", "purple", "orange", "brown", "pink", "black"))
}

dev.off()
```

```{r}
# for (i in seq(length(names(snfiles)))){
# 
#   corPlot(snfiles[[i]], pollutants = c("no2.ML", "no2", "co.ML", "co", "no.ML", "correctedNO",  "o3.ML", "o3"),y.relation = "free", main = paste0(names(snfiles)[i], "Correlation Plots"))
# }
```