---
title: "PM Filter"
---

In this notebook, we'll evaluate the PM data. We'll remove extreme, explained meteorological phenomena which eclipse a holistic story of PM presence at the sites. 

# Importing 

```{r, warning= FALSE, error=FALSE}
library(lubridate) #date and time functions 
library(data.table) #to use the data.table variable type
library(dplyr) #this library allows you to use the %>% operator
library(tidyr) #this library lets you use the complete function to account for time syncing
library(openair) #for plotting and analysis
library(readr)
library(ggplot2)
library(sjmisc)
library(plotly)
```

## Importing Files 

```{r}
nm <- list.files(path = "./", pattern = "-final-w-ML.csv", full.names = TRUE)
snfiles <-  sapply(nm,  FUN = function(x) fread(file=x,  data.table = TRUE), simplify = FALSE,USE.NAMES = TRUE)
```

```{r}
names(snfiles) <- lapply(names(snfiles), function(x) substr(x, 3, 6)) 
```


```{r}
set_vector <- function(snfile){
  snfile$date <- ymd_hms(snfile$date, tz = "UTC")
  snfile$date_local <- ymd_hms(snfile$date_local, tz = "America/New_York")
  return(snfile)
}
snfiles <- sapply(snfiles, function(x) set_vector(x), simplify = FALSE,USE.NAMES = TRUE)
```
## Importing Important Functions 

Here, we define a function that prints out the polar plots for the PM data.
```{r}
generate_PM_polarplots <- function(sn, snstring){
  #generates the polar plots for the variables we are interested in 
  polarPlot(sn, pollutant = "pm1.ML", main = paste0(snstring, " PM1.ML Polar Plot", sep= " "))
  polarPlot(sn, pollutant = "pm25.ML", main = paste0(snstring, " PM2.5.ML Polar Plot", sep= " "))
  polarPlot(sn, pollutant = "pm10.ML", main = paste0(snstring, " PM10.ML Polar Plot", sep= " "))

}
```

# Explore Eastern PM data 

In the polar plots of the PM data, there was an overwhelming easterly component. In this section, we'll filter out that component, and see if there's some temporal pattern between them. 

Based on the polar plots we can estimate what the elevated pollutant levels from the eastern source are. In the following section, we'll filter out the eastern source by wind direction, wind speed and pollutant level. 

We'll create a function that gathers the dates which the wind is coming from the East and exceeds 13 m/s (high wind/ gale on the Beaufort scale). We also filter those dates for PM values that exceed given values, which are the flag parameters into the function.

```{r}
flagFilter <- function(noreaster_data, flag1, flag2, flag3){
  
  
  PM1_date <- noreaster_data$date_local[which(noreaster_data$pm1.ML >= flag1 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] 
  
  PM25_date <- noreaster_data$date_local[which(noreaster_data$pm25.ML >= flag2 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] 
  
  PM10_date <- noreaster_data$date_local[which(noreaster_data$pm10.ML >= flag3 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] 
  
  noreaster_dates <- list("PM1 dates"= PM1_date, "PM 2.5 dates"= PM25_date, "PM10 dates"= PM10_date)
  
  return(noreaster_dates)
}
```

```{r}
sn45PMdates <- flagFilter(snfiles$sn45, 17, 17, 150)

sn49PMdates <- flagFilter(snfiles$sn49, 15, 17, 125)
sn62PMdates <- flagFilter(snfiles$sn62, 7, 9, 12)
sn67PMdates <- flagFilter(snfiles$sn67, 22, 22, 175)
sn72PMdates <- flagFilter(snfiles$sn72, 15, 17, 100)
```

We see that many of the dates listed in the datasets above took place in the mornings of 10/17/2019 and 1/16/2021. A quick google search shows that during these times there were storms with high (~40 mph) E winds. 

Since we know that these are isolated extreme weather events, we can filter them out. Let's try filtering the events themselves, filtering the characteristics that describe the event, and both. 

```{r}
rm(sn45PMdates, sn49PMdates, sn62PMdates, sn67PMdates, sn72PMdates)
```


## Filter by events 

Let's try filtering just the aforementioned dates from the PM data.


```{r}
eventFilter <- function(noreaster_data, flag1, flag2, flag3){
  
  noreaster_data$pm1.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  
    noreaster_data$pm1.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  
  return(noreaster_data)
}


snfiles_events_Test <- snfiles
snfiles_events_Test$sn45 <- eventFilter(snfiles$sn45, 17, 17, 150)

snfiles_events_Test$sn49 <- eventFilter(snfiles$sn49, 15, 17, 125)
snfiles_events_Test$sn62 <- eventFilter(snfiles$sn62, 7, 9, 12)
snfiles_events_Test$sn67 <- eventFilter(snfiles$sn67, 22, 22, 175)
snfiles_events_Test$sn72 <- eventFilter(snfiles$sn72, 15, 17, 100)

```

We can print out the polar plots of the resulting data 

```{r}
# pdf("PM-events.pdf")
for (i in seq(length(snfiles))){
 generate_PM_polarplots(snfiles_events_Test[[i]], names(snfiles)[i])
}
# dev.off()
```



## Filter by ws-wd-PMlevels

We can also try filtering the data the same way we found the storms: by filtering wind speed, wind direction and pollutant levels. We do this below 


```{r}
windFilter <- function(noreaster_data, flag1, flag2, flag3){
  
  noreaster_data$pm1.ML[which(noreaster_data$pm1.ML >= flag1 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$pm25.ML >= flag2 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$pm10.ML >= flag3 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  
  return(noreaster_data)
}


snfilesTest2 <- snfiles
snfilesTest2$sn45 <- windFilter(snfiles$sn45, 17, 17, 150)

snfilesTest2$sn49 <- windFilter(snfiles$sn49, 15, 17, 125)
snfilesTest2$sn62 <- windFilter(snfiles$sn62, 7, 9, 12)
snfilesTest2$sn67 <- windFilter(snfiles$sn67, 22, 22, 175)
snfilesTest2$sn72 <- windFilter(snfiles$sn72, 15, 17, 100)

# pdf("PM-ws-wd-filter.pdf")
for (i in seq(length(snfiles))){
 generate_PM_polarplots(snfilesTest2[[i]], names(snfiles)[i])
}
# dev.off()
```

```{r}
rm(snfilesTest2)
```


## Filter both 

What does the data look like if we combine the two filters above? 


```{r}
bothFilter <- function(noreaster_data, flag1, flag2, flag3){
  
  noreaster_data$pm1.ML[which(noreaster_data$pm1.ML >= flag1 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$pm25.ML >= flag2 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$pm10.ML >= flag3 & noreaster_data$wd > 45 & noreaster_data$wd < 140 & noreaster_data$ws > 13)] <- NA
  
  noreaster_data$pm1.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2019-10-16 22:00:00" & noreaster_data$date_local < "2019-10-17 12:00:00")] <- NA
  
    noreaster_data$pm1.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2021-01-16 00:00:00" & noreaster_data$date_local < "2021-01-16 12:00:00")] <- NA
  
  return(noreaster_data)
}
```

```{r}
snfilesTest3 <- snfiles
snfilesTest3$sn45 <- bothFilter(snfiles$sn45, 17, 17, 150)
snfilesTest3$sn49 <- bothFilter(snfiles$sn49, 15, 17, 125)
snfilesTest3$sn62 <- bothFilter(snfiles$sn62, 7, 9, 12)
snfilesTest3$sn67 <- bothFilter(snfiles$sn67, 22, 22, 175)
snfilesTest3$sn72 <- bothFilter(snfiles$sn72, 15, 17, 100)
```


```{r}
# pdf("PM-with-bothfilters.pdf")
# for (i in seq(length(snfiles))){
#  generate_PM_polarplots(snfilesTest3[[i]], names(snfiles)[i])
# }
# dev.off()
```

```{r}
rm(snfilesTest3)
```


# Second pass

After filtering out the two extreme weather events above, we see that there still appears to be a strong influence from the South East. In this section we will determine the source of it, and how to approach it. 

## Determining Temporal Pattern 

Like above, we will create a filter that will tell us when those elevated pollutant levels happen. 


```{r}
flagFilter2 <- function(noreaster_data, flag1, flag2, flag3){
  
  
  PM1_date <- noreaster_data$date_local[which(noreaster_data$pm1.ML >= flag1 & noreaster_data$wd > 135 & noreaster_data$wd < 190 & noreaster_data$ws > 12)] 
  
  PM25_date <- noreaster_data$date_local[which(noreaster_data$pm25.ML >= flag2 & noreaster_data$wd > 135 & noreaster_data$wd < 190 & noreaster_data$ws > 12)] 
  
  PM10_date <- noreaster_data$date_local[which(noreaster_data$pm10.ML >= flag3 & noreaster_data$wd > 135 & noreaster_data$wd < 190 & noreaster_data$ws > 12)] 
  
  noreaster_dates <- list("PM1 dates"= PM1_date, "PM 2.5 dates"= PM25_date, "PM10 dates"= PM10_date)
  
  return(noreaster_dates)
}
```

```{r}
# sn45PMdates <- flagFilter2(snfiles$sn45, 10, 10, 55)
# sn49PMdates <- flagFilter2(snfiles$sn49, 9, 10, 35)
# sn62PMdates <- flagFilter2(snfiles$sn62, 4, 4.5, 12)
# sn67PMdates <- flagFilter2(snfiles$sn67, 10, 10, 35)
```


### Analysis 

Now we see that events happening over multiple days are contributing to the elevated pollutant levels. Because of this, it no longer makes sense to filter by event. 

## Visual Inspection 

In this section we will describe two visual inspection methods we had wanted to employ, but which we ultimately did not end up using. 

We expected that, on a time series plot, nor'easter events appear as elevated PM levels for extended periods of time, accompanied by constant wind direction and high wind speed. We wanted to visually inspect the data for these events. 

### Plotly Visual Inspection

Technically, creating separated time series panels of the different PMs with the plotly interactive feature proved to be difficult. Therefore, we stopped pursuing this option. 

### Inspection of singular event 

Based on the flag filter above, we have a list of possible nor'easters. We will plot one such possible event below. 

## WS filter 

Ultimately, because the above two methods did not address our needs accurately, we have decided to opt for a filter that solely responds to wind speed. We will filter out PM data taken when wind speeds are over 13.5 m/s. According to the Beaufort scale, wind speeds over 13.5 m/s are classified as gale, and cause whole trees to move as well as "white foam from breaking waves begins to be blown in streaks along the direction of the wind". In other words, high windspeeds cause particulates to be blown into the air from the sea. This is especially relevant for us, since a number of our sites are very close to the ocean. Thus, we believe that this filter is justified. 

```{r}
wsfilter <- function(sensor){
  sensor$pm1.ML[which(sensor$ws > 13.5)] <- NA
  sensor$pm25.ML[which(sensor$ws > 13.5)] <- NA
  sensor$pm10.ML[which(sensor$ws > 13.5)] <- NA
  
  return(sensor)
}

snfiles_ws <- lapply(snfiles, function(x) wsfilter(x))
```

```{r}
# pdf("ws-filter-PMdata.pdf")
# for (i in seq(length(snfiles))){
#  generate_PM_polarplots(snfiles_ws[[i]], names(snfiles)[i])
# }
# dev.off()
```
## Adding event filter 

Since there still appears to be some Southeastern activity, we will also be reapplying our events filter, which we also believe is justified. 

```{r}
snfiles_ws_events_Test <- lapply(snfiles_events_Test, function(x) wsfilter(x))
```


```{r}
# pdf("ws-and-events-filter-PM.pdf")
# for (i in seq(length(snfiles))){
#  generate_PM_polarplots(snfiles_ws_events_Test[[i]], names(snfiles)[i])
# }
# dev.off()
```

# Third filter 

Another filter method we are trying out is the following: 

- filter out dates where PM 10 levels are elevated from 90-180 during times when ws > 13.5
- set the PM values for those dates to NA 

Since there is the same SE signature on all the sites, we still believe there is a noreaster event. If there is a noreaster event, then PM 10 is most effected, since PM 10 captures seaspray. This way, we can indirectly search for noreaster events 


```{r}
seasprayflagFilter <- function(noreaster_data, flag){
  
  
  
  PM10_date <- noreaster_data$date_local[which(noreaster_data$pm10.ML >= flag & noreaster_data$wd > 90 & noreaster_data$wd < 180 & noreaster_data$ws > 13.5)] 
  

  
  return(PM10_date)
}
```

```{r}
sn45PMdates <- seasprayflagFilter(snfiles_events_Test$sn45, 10)
sn46PMdates <- seasprayflagFilter(snfiles_events_Test$sn46, 30)
sn49PMdates <- seasprayflagFilter(snfiles_events_Test$sn49, 40)
sn62PMdates <- seasprayflagFilter(snfiles_events_Test$sn62, 15)
sn67PMdates <- seasprayflagFilter(snfiles_events_Test$sn67, 40)
sn72PMdates <- seasprayflagFilter(snfiles_events_Test$sn72, 40)

```

```{r}
incommon <- Reduce(intersect, list(sn45PMdates,  
                       
                       sn67PMdates, sn72PMdates))
```

```{r}
for(ind in incommon){
  print(sn45PMdates[which(sn45PMdates==ind)])
}
```


```{r}
rm(sn45PMdates, sn49PMdates, sn62PMdates, sn67PMdates, sn72PMdates)
```


```{r}
thirdFilter <- function(noreaster_data){
  
  
  noreaster_data$pm1.ML[which(noreaster_data$date_local > "2020-08-04 10:00:00" & noreaster_data$date_local < "2020-08-04 24:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local >"2020-08-04 10:00:00" & noreaster_data$date_local < "2020-08-04 24:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local >"2020-08-04 10:00:00" & noreaster_data$date_local < "2020-08-04 24:00:00")] <- NA
  
  noreaster_data$pm1.ML[which(noreaster_data$date_local > "2020-12-25 00:00:00" & noreaster_data$date_local < "2020-12-25 12:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2020-12-25 00:00:00" & noreaster_data$date_local < "2020-12-25 12:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2020-12-25 00:00:00" & noreaster_data$date_local < "2020-12-25 12:00:00")] <- NA
  
  noreaster_data$pm1.ML[which(noreaster_data$date_local > "2020-11-30 10:00:00" & noreaster_data$date_local < "2020-11-30 22:00:00")] <- NA
  noreaster_data$pm25.ML[which(noreaster_data$date_local > "2020-11-30 10:00:00" & noreaster_data$date_local < "2020-11-30 22:00:00")] <- NA
  noreaster_data$pm10.ML[which(noreaster_data$date_local > "2020-11-30 10:00:00" & noreaster_data$date_local < "2020-11-30 22:00:00")] <- NA
  
  return(noreaster_data)
}
```

```{r}
snfilesTest3 <- snfiles_events_Test
snfilesTest3$sn45 <- thirdFilter(snfiles_events_Test$sn45)
snfilesTest3$sn49 <- thirdFilter(snfiles_events_Test$sn49)
snfilesTest3$sn62 <- thirdFilter(snfiles_events_Test$sn62)
snfilesTest3$sn67 <- thirdFilter(snfiles_events_Test$sn67)
snfilesTest3$sn72 <- thirdFilter(snfiles_events_Test$sn72)
```


```{r}
# pdf("PM-with-thirdfilters.pdf")
# for (i in seq(length(snfiles))){
#  generate_PM_polarplots(snfilesTest3[[i]], names(snfiles)[i])
# }
# dev.off()
```

# Heatmap 

```{r}
# combined_list$dayOfMonth <- format(combined_list$date_local, format="%d")
```

```{r}
# trendLevel(combined_list, x = "month", y = "dayOfMonth", pollutant = "pm10.ML", type = "sn")
```
```{r}
# pdf("heatmap.pdf")
# trendLevel(subset(combined_list, sn == "sn45"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn45")
# trendLevel(subset(combined_list, sn == "sn46"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn46")
# trendLevel(subset(combined_list, sn == "sn49"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn49")
# trendLevel(subset(combined_list, sn == "sn62"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn62")
# trendLevel(subset(combined_list, sn == "sn67"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn67")
# trendLevel(subset(combined_list, sn == "sn72"), x = "month", y = "dayOfMonth", pollutant = "pm10.ML", main = "sn72")
# dev.off()
```



```{r}
# trendLevel(snfiles$sn45, pollutant = "pm10.ML", y = "weekday")
# trendLevel(snfiles$sn46, pollutant = "pm10.ML", y = "weekday")
# trendLevel(snfiles$sn49, pollutant = "pm10.ML", y = "weekday")
# trendLevel(snfiles$sn62, pollutant = "pm10.ML", y = "weekday")
# trendLevel(snfiles$sn67, pollutant = "pm10.ML", y = "weekday")
# trendLevel(snfiles$sn72, pollutant = "pm10.ML", y = "weekday")
```

# Export Cleaned PM data 

We can export the sensor data with the newly filtered PM. We can do so below: 

```{r}
mapply(
  fwrite, #apply function write table
  x=snfiles_ws_events_Test, file=paste(names(snfiles), "-final-w-ML-PM", ".csv", sep="")
)
```




