---
title: "R Notebook"
output: html_notebook
---

This is to see how the correlation are for all different devices in order to evaluate the possibility to use collected metereological data. Downloaded data from boston logan will be used as "measuring stock". 

Data are from 26th feb 2019 to 26th feb 2020

Use aggregate method instead of loops?



loading data and constructing data frames
```{r}
library(dplyr)
library(tidyverse) 
library(lubridate)
library(openair)

d45 <- read_csv("d45_190226-200226_raw.csv")
tz(d45$timestamp_local) <- "EST"

d46 <- read_csv("d46_190226-200226_raw.csv")
tz(d46$timestamp_local) <- "EST"

d49 <- read_csv("d49_190226-200226_raw.csv")
tz(d49$timestamp_local) <- "EST"

d62 <- read_csv("d62_190226-200226_raw.csv")
tz(d62$timestamp_local) <- "EST"

d67 <- read_csv("d67_190226-200226_raw.csv")
tz(d67$timestamp_local) <- "EST"

d72 <- read_csv("d72_190226-200226_raw.csv")
tz(d67$timestamp_local) <- "EST"


# pulling in BOS and formatting names
bos <- read_csv("bos_190226-200226_knot_dir_EDT.csv")

bos <- bos %>% select(c("valid",
                          "drct",
                          "sknt"))


names(bos)[names(bos) == "valid"] <- "timestamp_local"
names(bos)[names(bos) == "drct"] <- "wind_dir"
names(bos)[names(bos) == "sknt"] <- "wind_speed_kn"

# for some reason ( csv format? ) bos is reverse order, fixing that
bos$timestamp_local <- rev(bos$timestamp_local)
bos$wind_dir <- rev(bos$wind_dir)
bos$wind_speed <- rev(bos$wind_speed)
bos$wind_speed_ms <- rev(bos$wind_speed_ms)


# for some reason the wind dir is of charachter type
bos$wind_dir <- as.numeric(bos$wind_dir)

# adding wind speed vector in m/s
bos$wind_speed_ms = bos$wind_speed_kn*0.5144444

# appropriate timezone
tz(bos$timestamp_local) <- "EST"

```

Cleaning data, removing weird wind directions and too high wind values
```{r}
#d45

d45$wind_dir <- invisible(replace(d45$wind_dir, d45$wind_dir > 360,  NA)) 

d45$wind_dir <- invisible(replace(d45$wind_dir, d45$wind_dir > 360,  NA))

d45$wind_speed <- invisible(replace(d45$wind_speed,
                                      d45$wind_speed > 30,
                                      0))
                              
d45$wind_speed <- invisible(replace(d45$wind_speed,
                                    d45$wind_speed < 0,
                                    0))
#d46

d46$wind_dir <- invisible(replace(d46$wind_dir, d46$wind_dir > 360,  NA)) 

d46$wind_dir <- invisible(replace(d46$wind_dir, d46$wind_dir > 360,  NA))

d46$wind_speed <- invisible(replace(d46$wind_speed,
                                      d46$wind_speed > 30,
                                      0))
                              
d46$wind_speed <- invisible(replace(d46$wind_speed,
                                    d46$wind_speed < 0,
                                    0))
#d49

d49$wind_dir <- invisible(replace(d49$wind_dir, d49$wind_dir > 360,  NA)) 

d49$wind_dir <- invisible(replace(d49$wind_dir, d49$wind_dir > 360,  NA))

d49$wind_speed <- invisible(replace(d49$wind_speed,
                                      d49$wind_speed > 30,
                                      0))
                              
d49$wind_speed <- invisible(replace(d49$wind_speed,
                                    d49$wind_speed < 0,
                                    0))

#d62

d62$wind_dir <- invisible(replace(d62$wind_dir, d62$wind_dir > 360,  NA)) 

d62$wind_dir <- invisible(replace(d62$wind_dir, d62$wind_dir > 360,  NA))

d62$wind_speed <- invisible(replace(d62$wind_speed,
                                      d62$wind_speed > 30,
                                      0))
                              
d62$wind_speed <- invisible(replace(d62$wind_speed,
                                    d62$wind_speed < 0,
                                    0))
#d67

d67$wind_dir <- invisible(replace(d67$wind_dir, d67$wind_dir > 360,  NA)) 

d67$wind_dir <- invisible(replace(d67$wind_dir, d67$wind_dir > 360,  NA))

d67$wind_speed <- invisible(replace(d67$wind_speed,
                                      d67$wind_speed > 30,
                                      0))
                              
d67$wind_speed <- invisible(replace(d67$wind_speed,
                                    d67$wind_speed < 0,
                                    0))
#d72

d72$wind_dir <- invisible(replace(d72$wind_dir, d72$wind_dir > 360,  NA)) 

d72$wind_dir <- invisible(replace(d72$wind_dir, d72$wind_dir > 360,  NA))

d72$wind_speed <- invisible(replace(d72$wind_speed,
                                      d72$wind_speed > 30,
                                      0))
                              
d72$wind_speed <- invisible(replace(d72$wind_speed,
                                    d72$wind_speed < 0,
                                    0))

# bos
bos$wind_speed_ms <- invisible(replace(bos$wind_speed_ms,
                                      bos$wind_speed_ms > 30,
                                      0))
                              
bos$wind_speed_ms <- invisible(replace(bos$wind_speed_ms,
                                     bos$wind_speed_ms < 0,
                                      0))

# removing wind directions that does not make sense, bos
bos$wind_dir <- invisible(replace(bos$wind_dir, bos$wind_dir > 360, NA))
```


Creating rounded timestamp vectors
```{r}
bos$timestamp_local_min <- round_date(bos$timestamp_local,  
                                       unit = "minute")

d45$timestamp_local_min <- round_date(d45$timestamp_local, 
                                       unit = "minute")
d46$timestamp_local_min <- round_date(d46$timestamp_local, 
                                       unit = "minute")
d49$timestamp_local_min <- round_date(d49$timestamp_local, 
                                       unit = "minute")
d62$timestamp_local_min <- round_date(d62$timestamp_local, 
                                       unit = "minute")
d67$timestamp_local_min <- round_date(d67$timestamp_local, 
                                       unit = "minute")
d72$timestamp_local_min <- round_date(d72$timestamp_local, 
                                       unit = "minute")
```


Removing all values not on 5 min interval & creating time range in bos
```{r}
saveVec <- seq(0, 55, by = 5) # used to save values mathing these
timestamp_local_min <- bos$timestamp_local_min[minute(bos$timestamp_local_min)%in%saveVec]

mergedf <- data.frame(timestamp_local_min)
bos <- merge(mergedf,
              bos,
              by = "timestamp_local_min")


bos$start <- bos$timestamp_local_min-minutes(2)
bos$end <- bos$timestamp_local_min+minutes(2)
bos$range <- bos$start%--%bos$end
```


Creating hourly mean values for bos
```{r}
bos$timestamp_local_h <- round_date(bos$timestamp_local_min, unit = "hour")
hVec <- seq(from = bos$timestamp_local_min[1],
            to = tail(bos$timestamp_local_min,1),
            by = "hour")


bos.wd.hmeans <- NULL
bos.ws.hmeans <- NULL 
for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- bos$timestamp_local_h == i
   
    # indices of true elements of above 
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- bos$wind_speed_ms[xind]   
    
    # vector with wd elements that matched range
    wdVec <- bos$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    bos.ws.hmeans <- append(bos.ws.hmeans, mean(wsVec))
    bos.wd.hmeans <- append(bos.wd.hmeans, mean(wdVec))
}

```



creating dataframe with d45 and bos on 5 min intervals
```{r}
wd.means <- (NULL)
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) {
    
    # boolean vector for times within rane
    boolVec <- d45$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d45$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d45$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec))
}

df45 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms,
                  bos$wind_dir,
                  ws.means,
                  wd.means)




```

creating dataframe with d45 and bos on 1 hour intervals
```{r}

d45$timestamp_local_h <- round_date(d45$timestamp_local_min, unit = "hour")

d45.wd.hmeans <- NULL
d45.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d45$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d45$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d45$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d45.ws.hmeans <- append(d45.ws.hmeans, mean(wsVec))
    d45.wd.hmeans <- append(d45.wd.hmeans, mean(wdVec))
}

hdf45 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans,
                   d45.wd.hmeans,
                   d45.ws.hmeans)
```



d46 and bos on 5 min intervals
```{r}
wd.means <- (NULL)
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) {
     
    # boolean vector for times within rane   
    boolVec <- d46$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d46$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d46$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec))
}

df46 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms,
                  bos$wind_dir,
                  ws.means,
                  wd.means)
```

d46 and bos on 1 hour intervals
```{r}

d46$timestamp_local_h <- round_date(d46$timestamp_local_min, unit = "hour")

d46.wd.hmeans <- NULL
d46.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d46$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d46$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d46$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d46.ws.hmeans <- append(d46.ws.hmeans, mean(wsVec))
    d46.wd.hmeans <- append(d46.wd.hmeans, mean(wdVec))
}

hdf46 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans,
                   d46.wd.hmeans,
                   d46.ws.hmeans)
```


d49 and bos on 5 min intervals
```{r}
wd.means <- (NULL)
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) {
    
    # boolean vector for times within rane
    boolVec <- d49$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d49$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d49$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec))
}

df49 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms,
                  bos$wind_dir,
                  ws.means,
                  wd.means)
```

d49 and bos on 1 hour intervalsls
```{r}

d49$timestamp_local_h <- round_date(d49$timestamp_local_min, unit = "hour")

d49.wd.hmeans <- NULL
d49.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d49$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d49$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d49$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d49.ws.hmeans <- append(d49.ws.hmeans, mean(wsVec))
    d49.wd.hmeans <- append(d49.wd.hmeans, mean(wdVec))
}

hdf49 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans,
                   d49.wd.hmeans,
                   d49.ws.hmeans)
```
 


d62 and bos on 5 min intervals
```{r}
wd.means <- (NULL)
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) {
    
    # boolean vector for times within rane
    boolVec <- d62$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current  
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d62$wind_speed[xind]   
    
    # vector with wd elements that matched range 
    wdVec <- d62$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec)) 
} 

df62 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms,
                  bos$wind_dir,
                  ws.means,
                  wd.means) 
``` 

d62 and bos on 1 hour intervals
```{r}

d62$timestamp_local_h <- round_date(d62$timestamp_local_min, unit = "hour")

d62.wd.hmeans <- NULL
d62.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d62$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d62$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d62$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d62.ws.hmeans <- append(d62.ws.hmeans, mean(wsVec))
    d62.wd.hmeans <- append(d62.wd.hmeans, mean(wdVec))
}

hdf62 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans,
                   d62.wd.hmeans,
                   d62.ws.hmeans)
```


d67 and bos on 5 min intervals
```{r}
wd.means <- (NULL)
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) {
    
    # boolean vector for times within rane
    boolVec <- d67$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d67$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d67$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec))
}

df67 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms,
                  bos$wind_dir,
                  ws.means,  
                  wd.means) 
```

d67 and bos on 1 hour intervals
```{r}

d67$timestamp_local_h <- round_date(d67$timestamp_local_min, unit = "hour")

d67.wd.hmeans <- NULL
d67.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d67$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d67$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d67$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d67.ws.hmeans <- append(d67.ws.hmeans, mean(wsVec))
    d67.wd.hmeans <- append(d67.wd.hmeans, mean(wdVec))
}

hdf67 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans,
                   d67.wd.hmeans,
                   d67.ws.hmeans)
```



d72 and bos on 5 min intervals
```{r}
wd.means <- (NULL) 
ws.means <- (NULL)
for (i in (1:length(bos$range)) ) { 
    
    # boolean vector for times within rane
    boolVec <- d72$timestamp_local_min%within%bos$range[i]
    if (sum(boolVec == TRUE) > 8)  {
        print(paste("+ 8 matches Where i =  ", i))
    }
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    if (i > 1) { # checking for double counting
        d.match <- xind%in%xind.last
        if (TRUE%in%d.match) {
            print(paste("double match on indices:",
                        xind[which(d.match, TRUE)]))
        }
    }
    xind.last <- xind # from last loop
    
    
    # vector with ws elements that matched range
    wsVec <- d72$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d72$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    ws.means <- append(ws.means, mean(wsVec))
    wd.means <- append(wd.means, mean(wdVec))
}

df72 <- data.frame(bos$timestamp_local_min,
                  bos$wind_speed_ms, 
                  bos$wind_dir,
                  ws.means, 
                  wd.means)
```

d72 and bos on 1 hour intervals
```{r}

d72$timestamp_local_h <- round_date(d72$timestamp_local_min, unit = "hour")

d72.wd.hmeans <- NULL
d72.ws.hmeans <- NULL

for (i in hVec) {
    
    # boolean vector for times within rane
    boolVec <- d72$timestamp_local_h == i
   
    # indices of true elements of above
    xind <- which(boolVec == TRUE) # current 
    #if (length(xind > 15)){print(paste("more than 15 values at:", i))}
    
    # vector with ws elements that matched range
    wsVec <- d72$wind_speed[xind]   
    
    # vector with wd elements that matched range
    wdVec <- d72$wind_dir[xind]
    
    # concatenating vectors with means with mean of current matches
    d72.ws.hmeans <- append(d72.ws.hmeans, mean(wsVec))
    d72.wd.hmeans <- append(d72.wd.hmeans, mean(wdVec))
}

hdf72 <- data.frame(hVec,
                   bos.ws.hmeans,
                   bos.wd.hmeans, 
                   d72.wd.hmeans, 
                   d72.ws.hmeans) 
```


Correlation analysis df45 & hdf45
```{r}
# 5 min intervals scatterplot and corplot

plot(df45$ws.means,
     df45$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df45$wd.means,
     df45$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df45)

# 1 hour intervals scatterplot and corplot

plot(hdf45$d45.ws.hmeans,
     hdf45$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf45$d45.wd.hmeans,
     hdf45$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf45)

```


Correlation analysis df46 & hdf46
```{r}
# 5 min intervals scatterplot and corplot

plot(df46$ws.means,
     df46$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df46$wd.means,
     df46$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df46)

# 1 hour intervals scatterplot and corplot

plot(hdf46$d46.ws.hmeans,
     hdf46$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf46$d46.wd.hmeans,
     hdf46$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf46)

```


Correlation analysis df49 & hdf49
```{r}
# 5 min intervals scatterplot and corplot

plot(df49$ws.means,
     df49$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df49$wd.means,
     df49$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df49)

# 1 hour intervals scatterplot and corplot

plot(hdf49$d49.ws.hmeans,
     hdf49$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf49$d49.wd.hmeans,
     hdf49$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf49)

```


Correlation analysis df62 & hdf62
```{r}
# 5 min intervals scatterplot and corplot

plot(df62$ws.means,
     df62$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df62$wd.means,
     df62$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df62)

# 1 hour intervals scatterplot and corplot

plot(hdf62$d62.ws.hmeans,
     hdf62$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf62$d62.wd.hmeans,
     hdf62$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf62)

```


Correlation analysis df67 & hdf67
```{r}
# 5 min intervals scatterplot and corplot

plot(df67$ws.means,
     df67$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df67$wd.means,
     df67$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df67)

# 1 hour intervals scatterplot and corplot

plot(hdf67$d67.ws.hmeans,
     hdf67$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf67$d67.wd.hmeans,
     hdf67$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf67)

```


Correlation analysis df72 & hdf72
```{r}
# 5 min intervals scatterplot and corplot

plot(df72$ws.means,
     df72$bos.wind_speed_ms,
     xlim = c(0,35), ylim = c(0, 35))

plot(df72$wd.means,
     df72$bos.wind_dir,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(df72)

# 1 hour intervals scatterplot and corplot

plot(hdf72$d72.ws.hmeans,
     hdf72$bos.ws.hmeans,
     xlim = c(0,35), ylim = c(0, 35))

plot(hdf72$d72.wd.hmeans,
     hdf72$bos.wd.hmeans,
     xlim = c(0,400), ylim = c(0, 400))

corPlot(hdf72)

```


Monthly means with std, bos
```{r}




wsmeans.bos <- aggregate(bos["wind_speed_ms"],
                       format(bos["timestamp_local"],"%Y-%m"),
                       mean,
                       na.rm = TRUE)
ws.sd.bos <- aggregate(bos["wind_speed_ms"],
                       format(bos["timestamp_local"],"%Y-%m"),
                       sd,
                       na.rm = TRUE)
wdmeans.bos <- aggregate(bos["wind_dir"],
                       format(bos["timestamp_local"],"%Y-%m"),
                       mean, 
                       na.rm = TRUE) 
wd.sd.bos <- aggregate(bos["wind_dir"],
                       format(bos["timestamp_local"],"%Y-%m"),
                       sd,
                       na.rm = TRUE)

bos.monthly <- data.frame(wsmeans.bos$timestamp_local,
                          wsmeans.bos$wind_speed_ms,
                          ws.sd.bos$wind_speed_ms,
                          wdmeans.bos$wind_dir,
                          wd.sd.bos$wind_dir)

names(bos.monthly)[names(bos.monthly) == "wsmeans.bos.timestamp_local"] <- "timestamp_local"
names(bos.monthly)[names(bos.monthly) == "wsmeans.bos.wind_speed_ms"] <- "bos.ws"
names(bos.monthly)[names(bos.monthly) == "ws.sd.bos.wind_speed_ms"] <- "bos.sd.ws"
names(bos.monthly)[names(bos.monthly) == "wdmeans.bos.wind_dir"  ] <- "bos.wd"
names(bos.monthly)[names(bos.monthly) == "wd.sd.bos.wind_dir"] <- "bos.sd.wd"

```


Monthly means with std, d45
```{r}

d45.timestamp_local_month <- as.POSIXct(d45$timestamp_local,
                                        format = "%Y-%m")

d45$timestamp_local <- as.Date(d45$timestamp_local)

d45$month <- months(d45$timestamp_local)

d45$Year <- format(d45$timestamp_local,format="%y")


ws.d45 <- aggregate(wind_speed ~ month + Year,
                                 d45,
                                 mean )

wd.d45 <- aggregate(wind_speed ~ month + Year,
                                 d45,
                                 mean )

ws.sd.d45 <- aggregate(wind_speed ~ month + Year,
                                 d45,
                                 sd )

wd.sd.d45 <- aggregate(wind_speed ~ month + Year,
                                 d45,
                                 sd )

ws.d45 <- ws.d45 %>% arrange(match(month, month.name))
ws.d45 <- ws.d45 %>% arrange(Year)

wd.d45 <- wd.d45 %>% arrange(match(month, month.name))
wd.d45 <- wd.d45 %>% arrange(Year)

ws.sd.d45 <- ws.sd.d45 %>% arrange(match(month, month.name))
ws.sd.d45 <- ws.sd.d45 %>% arrange(Year)

wd.sd.d45 <- wd.sd.d45 %>% arrange(match(month, month.name))
wd.sd.d45 <- wd.sd.d45 %>% arrange(Year)




```

Monthly means with std, d46
```{r}

d46$timestamp_local <- as.Date(d46$timestamp_local)

d46$month <- months(d46$timestamp_local)

d46$Year <- format(d46$timestamp_local,format="%y")

ws.d46 <- aggregate(wind_speed ~ month + Year ,
                                 d46,
                                 mean )

wd.d46 <- aggregate(wind_speed ~ month + Year ,
                                 d46,
                                 mean )

ws.sd.d46 <- aggregate(wind_speed ~ month + Year ,
                                 d46,
                                 sd )

wd.sd.d46 <- aggregate(wind_speed ~ month + Year , 
                                 d46,
                                 sd )

ws.d46 <- ws.d46 %>% arrange(match(month, month.name))
ws.d46 <- ws.d46 %>% arrange(Year)

wd.d46 <- wd.d46 %>% arrange(match(month, month.name))
wd.d46 <- wd.d46 %>% arrange(Year)

ws.sd.d46 <- ws.sd.d46 %>% arrange(match(month, month.name))
ws.sd.d46 <- ws.sd.d46 %>% arrange(Year)

wd.sd.d46 <- wd.sd.d46 %>% arrange(match(month, month.name))
wd.sd.d46 <- wd.sd.d46 %>% arrange(Year)
```

Monthly means with std, d49
```{r}
 
d49$timestamp_local <- as.Date(d49$timestamp_local)

d49$month <- months(d49$timestamp_local)

d49$Year <- format(d49$timestamp_local,format="%y")

ws.d49 <- aggregate(wind_speed ~ month + Year ,
                                 d49,
                                 mean )

wd.d49 <- aggregate(wind_speed ~ month + Year ,
                                 d49,
                                 mean )

ws.sd.d49 <- aggregate(wind_speed ~ month + Year ,
                                 d49,
                                 sd )


wd.sd.d49 <- aggregate(wind_speed ~ month + Year , 
                                 d49,
                                 sd )

ws.d49 <- ws.d49 %>% arrange(match(month, month.name))
ws.d49 <- ws.d49 %>% arrange(Year)

wd.d49 <- wd.d49 %>% arrange(match(month, month.name))
wd.d49 <- wd.d49 %>% arrange(Year)

ws.sd.d49 <- ws.sd.d49 %>% arrange(match(month, month.name))
ws.sd.d49 <- ws.sd.d49 %>% arrange(Year)

wd.sd.d49 <- wd.sd.d49 %>% arrange(match(month, month.name))
wd.sd.d49 <- wd.sd.d49 %>% arrange(Year)
```

Monthly means with std, d62
```{r}
 
d62$timestamp_local <- as.Date(d62$timestamp_local)

d62$month <- months(d62$timestamp_local)

d62$Year <- format(d62$timestamp_local,format="%y")

ws.d62 <- aggregate(wind_speed ~ month + Year ,
                                 d62,
                                 mean )

wd.d62 <- aggregate(wind_speed ~ month + Year ,
                                 d62,
                                 mean )

ws.sd.d62 <- aggregate(wind_speed ~ month + Year ,
                                 d62,
                                 sd )

wd.sd.d62 <- aggregate(wind_speed ~ month + Year ,
                                 d62,
                                 sd )

ws.d62 <- ws.d62 %>% arrange(match(month, month.name))
ws.d62 <- ws.d62 %>% arrange(Year)

wd.d62 <- wd.d62 %>% arrange(match(month, month.name))
wd.d62 <- wd.d62 %>% arrange(Year)

ws.sd.d62 <- ws.sd.d62 %>% arrange(match(month, month.name))
ws.sd.d62 <- ws.sd.d62 %>% arrange(Year)

wd.sd.d62 <- wd.sd.d62 %>% arrange(match(month, month.name))
wd.sd.d62 <- wd.sd.d62 %>% arrange(Year)
```

Monthly means with std, d67
```{r}
 
d67$timestamp_local <- as.Date(d67$timestamp_local)

d67$month <- months(d67$timestamp_local)

d67$Year <- format(d67$timestamp_local,format="%y")

ws.d67 <- aggregate(wind_speed ~ month + Year ,
                                 d67,
                                 mean )

wd.d67 <- aggregate(wind_speed ~ month + Year ,
                                 d67,
                                 mean )

ws.sd.d67 <- aggregate(wind_speed ~ month + Year ,
                                 d67,
                                 sd )

wd.sd.d67 <- aggregate(wind_speed ~ month + Year ,
                                 d67,
                                 sd )

ws.d67 <- ws.d67 %>% arrange(match(month, month.name))
ws.d67 <- ws.d67 %>% arrange(Year)

wd.d67 <- wd.d67 %>% arrange(match(month, month.name))
wd.d67 <- wd.d67 %>% arrange(Year)

ws.sd.d67 <- ws.sd.d67 %>% arrange(match(month, month.name))
ws.sd.d67 <- ws.sd.d67 %>% arrange(Year)

wd.sd.d67 <- wd.sd.d67 %>% arrange(match(month, month.name))
wd.sd.d67 <- wd.sd.d67 %>% arrange(Year)
```

Monthly means with std, d72
```{r}

d72$timestamp_local <- as.Date(d72$timestamp_local)
 
d72$month <- months(d72$timestamp_local)
 
d72$Year <- format(d72$timestamp_local,format="%y")

ws.d72 <- aggregate(wind_speed ~ month + Year ,
                                 d72,
                                 mean )

wd.d72 <- aggregate(wind_dir ~ month + Year ,
                                 d72,
                                 mean )


ws.sd.d72 <- aggregate(wind_speed ~ month + Year ,
                                 d72,
                                 sd )
names(ws.sd.d72)[names(ws.sd.d72) == "wind_speed"] <- "wind_speed_sd"

wd.sd.d72 <- aggregate(wind_dir ~ month + Year ,
                                 d72,
                                 sd )
names(wd.sd.d72)[names(wd.sd.d72) == "wind_dir"] <- "wind_dir_sd"

monthly.d72 <- merge(
                    merge(
                        merge(ws.d72,
                              ws.sd.d72wd.d72),
                              ws.sd.d72),
                              wd.sd.d72)

monthly.d72 <- monthly.d72 %>% arrange(match(month, month.name))
monthly.d72 <- monthly.d72 %>% arrange(Year)





```