---
title: "IS Deliverable 6"
output:   
  md_document:
    variant: markdown_github
---

### Setup and Preprocessing 
#### Including Necessary Libraries

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(gridExtra)
library(grid)
library(openair)
library(data.table)
library(smps)
```

#### Importing Data and Setting Up Workspace
```{r}
#setting workspace
#setwd(choose.dir())
#next we import the data 
sn62 = read.csv("is2.csv", header = TRUE)

#this time I'm inporting data up until today as well
sn62_full = read.csv("IS_full.csv", header=TRUE)
```
#### Formatting 
```{r}
#changing the datetime format 
sn62$timestamp <- ymd_hms(sn62$timestamp)
#renaming the datasets, because openair needs a "date" column
#also setting wind_speed and wind_direction to ws and wd, b/c of wind rose function 
sn62 <- sn62 %>%
  rename(date = timestamp, ws = wind_speed, wd = wind_dir) 
#set the date column as the first column in the dataframe, for openair formatting
sn62[1] <- sn62$date

```
```{r}
#doing the same for the full dataset
sn62_full$timestamp <- ymd_hms(sn62_full$timestamp)
sn62_full <- sn62_full %>%
  rename(date = timestamp, ws = wind_speed, wd = wind_dir) 
sn62_full[1] <- sn62_full$date
```
#### Filtering 
```{r}
sn62 <- dplyr::filter(sn62, between(sn62$wd, 0, 360))
sn62 <-dplyr::filter(sn62, between(sn62$ws, 0, 30))
sn62 <-dplyr::filter(sn62, between(sn62$pressure, 98000, 120000))
sn62 <- dplyr::filter(sn62, between(sn62$temp_manifold, -10, 40))

```
```{r}
sn62_full <- dplyr::filter(sn62_full, between(sn62_full$wd, 0, 360))
sn62_full <-dplyr::filter(sn62_full, between(sn62_full$ws, 0, 30))
sn62_full <-dplyr::filter(sn62_full, between(sn62_full$pressure, 98000, 120000))
sn62_full <- dplyr::filter(sn62_full, between(sn62_full$temp_manifold, -10, 40))

```

# Spectrograms

First I got the smps library that Scott mentioned 
I had trouble installing it at first, but I followed this advice (I had the exact same error) https://github.com/rstudio/distill/issues/47

It looks like the data has to have time as its first column and then some sort of bin number for all the rest of the columns, this could be midpoint . or it should be diameter, so we can do midpoint like scott said 

so 
step 1 -> make a new dataframe in the right format
step 2-> run smps package

```{r}

# put air data in the right format
# mydata <- subset(sn62, select=c(3,26:40))
# class(sn62[3]) 
# mydata <- mydata %>% mutate(date = ymd_hms(date)) #convert Time to a timestamp
# setnames(mydata, old =c('bin0','bin1','bin2','bin3','bin4','bin5','bin6','bin7','bin8','bin9','bin10','bin11','bin12','bin13', 'bin14'), new = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14"))
# prepared_data <- prepare_data(mydata)


```

This isn't working yet, but I'll work on it next week !

# Remaking PollutionRose

```{r}
breaks2 = c(0,(1/1000)*max(sn62$bin0),(1/100)*max(sn62$bin0), (1/10)*max(sn62$bin0), max(sn62$bin0))

pollutionRose(sn62, pollutant  = "bin0",breaks=breaks2)
```
I think that I finally like this pollutionRose plot. Although it still has the bulk of the data in the [1,9] range, I can better see how much that is to larger and smaller values. 

Another thing, however, was that I was expecting high values from the Northeast. Are they there, but I can't see them? Are they just small, comparatively to the other data?

I feel like I still don't understand this plot well. Are the circles representing what percentage of the data is facing that direction? So is 20% of the largest bin facing West?


# Plotting Bin0 vs co_we 

This is a continuation of the work from Deliverable 5

```{r}
ggplot(sn62,aes(x=co_we,y=bin0)) + geom_point(alpha = 0.01)
```
I tried getting a better sense of the data, by making a small alpha value (ie , making the data points more transparent), however, it still seems very dense. 

```{r}
ggplot(sn62,aes(x=co_we,y=bin0)) + stat_binhex()
```
I feel like this graph is more telling. It looks like the greatest concentration of bin0 data points happens when bin0 values and co_we values are low. At about 500 (I've forgotten the unit, but I'll check it for next week!) co_we units is the greatest number of data points, and it spans all possible bin0 values. On the other hand, bin0 values seem constant after 1000 co_we units. I feel like there's another part to this graph that I need to know to make sense of this data. Is anything after 1000 co_we unreliable data? Is 500 co_we units the most common co concentration? Is it associated with planes?
