---
title: "IS Deliverable 4"
output:
  md_document:
    variant: markdown_github
---
# Setup 

### Including Necessary Libraries

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(gridExtra)
library(grid)
library(openair)
```
#### Importing Data and Setting Up Workspace
```{r}
#setting workspace
#setwd(choose.dir())
#next we import the data 
sn62 = read.csv("is2.csv", header = TRUE)
```
#### Formatting 
```{r}
#changing the datetime format 
sn62$timestamp <- ymd_hms(sn62$timestamp)

#set the date column as the first column in the dataframe, for openair formatting
sn62[1] <- sn62$timestamp
#renaming the datasets, because openair needs a "date" column
#also setting wind_speed and wind_direction to ws and wd, b/c of wind rose function 
sn62 <- sn62 %>%
  rename(date = X, ws = wind_speed, wd = wind_dir) 
```
#### Filtering 
```{r}
sn62 <- dplyr::filter(sn62, between(sn62$wd, 0, 360))
sn62 <-dplyr::filter(sn62, between(sn62$ws, 0, 30))
sn62 <-dplyr::filter(sn62, between(sn62$pressure, 98000, 120000))
sn62 <- dplyr::filter(sn62, between(sn62$temp_manifold, -10, 40))

```
# Defining "normal" for bin data using statistical analysis

First, I look at a variety of sources

### Source 1 : DEcomposing into seasonal data 
The first [source](https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html) I read talked about splitting a timeseries dataset into its seasonal components. This could be one way of defining what normal bin data could be, with more accuracy. 

### Source 2: Time Series Characteristics and practices
This [source](https://online.stat.psu.edu/stat510/lesson/1/1.1) describes the main benefits of a timeseries analysis: to describe the important features, explain future events, etc. There are two basic types of time domain models. First are Autoregressive Integrated Moving Average), which relates present values to past values. The other is ordinary regression models, which can be helpful for an initial description of the data and doing simple forecasting. 

One really helpful section was the important characteristics to consider when looking at a time series graph. I hope to apply these questions to my next analysis. 

### Source 3: Time Series Analysis Online TextBook 
One source (which I forgot to cite and have now forgotten) describes a lot of things about time series analysis. In the first few chapters, there are a variety of time series analysis tools described, including : 
1. moving average 
2. Random Walk Model 
3. Autoregression 

### Other: 

Here's some helpful terms defined. 
stationary - time series whose properties arent time dependent. (ie seasonal trends arent stationary)

least squares estimation model 

## Testing Things Out 
```{r}
## calculate monthly means
monthly <- timeAverage(sn62, avg.time = "month")
seasonaly <- timeAverage(sn62, avg.time = "quarter")
```
```{r}
timePlot(monthly, pollutant= c("bin7", "bin10", "bin14"))
timePlot(seasonaly, pollutant= c("bin7", "bin10", "bin14"))
timePlot(sn62, pollutant= c("bin7", "bin10", "bin14"))
```
From this, I believe that seasonally is too broad to average.


## Changing bin size and reformatting
I found this [source](https://rpubs.com/NateByers/Openair), which uses "breaks" to determine bin size. In the example, they predefine the breakpoints, but I was thinking of doing max_val * (1/3), max_val * (2/3) , max_val. Or something similar 

Similarly, to Windrose I can add the parameter key.footer = "something" in order to label the graph more precisely. 